---
title: "Low-Rank Simulation"
output: html_document
params:
  B: 1000 # number of bootstraps
  N: 1000 # number of samples
  D: 100 # dimensionality
  K: 2 # latent dimensionality
---

This script generates the figures appearing in Section 3.1, comparing the three
bootstrap approaches on a permuted low-rank model. First, we load necessary
libraries.

```{r}
set.seed(1234)
library(expm)
library(irlba)
library(tidyverse)
library(MSLF)
theme_set(min_theme())
attach(params)
```

The block below simulates one realization from the mechanism in equations 4 - 7.
$X$ is a random low rank matrix observed with noise, and $y$ is a linear
combination of the latent coordinates from each row of $X$. In a more realistic
application, each row of $X$ would be an image, text, or audio sample, for
example, and we would use a feature extractor to convert it into a matrix. In
this simplified setup, though, we simple assume $X$ is a matrix.

```{r}
U <- r_ortho(N, D)
V <- r_ortho(D, D)
Sigma <- c(rep(100, 2), rep(0.0, D - K))
X <- U %*% diag(Sigma) %*% t(V) + rmat(N, D, 0.1)
beta <- c(c(1, -1), rep(0, D - K))
y <- U %*% diag(sqrt(Sigma)) %*% beta + rnorm(N, 0, .1)
```

Here is the strategy that extracts the features once and then applies the
parametric bootstrap of Section 2.2. The function $f$ simulates the feature
extraction step in equation 8 -- each call fo `boot_fun()` samples the truncated
SVD coordinates with noise and randomly permutes the columns. The
`align_to_list` function applies a Procrustes rotation to the
dimensionality-reduced learned features, and `align_with_truth` applies a second
rotation to align coordinates with the known coordinates from the previous
block. This last step is just so that we can overlay the truth on the estimated
coordinates, it would not be possible in a real application.

```{r}
f <- features(X)
boot_fun <- param_boot(f(X), K)
ud_combined <- list()
ud_combined[["parametric"]] <- rerun(B, boot_fun()$ub) %>%
  align_to_list() %>%
  align_with_truth(U, Sigma) %>%
  left_join(data.frame(i = seq_along(y), y = y))
```

The block below instead applies the nonparametric bootstrap of Section 2.1. It
computes $B$ feature extractors by calling `f` within the `rerun` loop (contrast
with th eblock above, which ran `f` once). The Procrustes analysis steps are the
same as above.

```{r}
f <- features(X)
Zb <- rerun(B, {
  full_features <- f(X)
  svf <- irlba(full_features, nv=K) 
  svf$u %*% diag(svf$d)
})
ud_combined[["nonparametric"]] <- align_to_list(Zb)  %>%
  align_with_truth(U, Sigma) %>%
  left_join(data.frame(i = seq_along(y), y = y))
```

The step below gives the compromise bootstrap of Section 2.3. It offers a of
compromise between repeated feature learning in the nonparametric bootstrap and
the single run from the parametric bootstrap. The code is also a mix of the two
approaches -- the first `rerun` step trains 100 feature extractors (not 1000
like before) and the second increases the total number of bootstraps to 1000 by
simulating from the parametric model.

```{r}
f <- features(X)
Zb <- rerun(100, f(X))
boot_fun <- param_boot_cmp(Zb)
ud_combined[["compromise"]] <- rerun(B, boot_fun()$ub) %>%
  align_to_list() %>%
  align_with_truth(U, Sigma) %>%
  left_join(data.frame(i = seq_along(y), y = y)) # adds a new list element with the aligned truth
```

Finally, we can compute the coverage of the confidence areas from each approach.
We are using `b == max(b)` to extract the true coordinates, since in the
`ud_combined` objects, the last element of each list is the aligned truth. It
seems that the compromise approach is conservative, the parametric approach is
potentially conservative, and the nonparametric approach is nearly exactly the
expected coverage.

```{r}
map(ud_combined, ~ {
  coverage(filter(., b < max(b)), filter(., b == max(b))) %>%
    mean()
})
```

The block below generates Figure 2. The `plot_overlay_truth` function is in
`R/vis.R` in this compendium. We are showing only the first 100 samples, since
showing all 1000 clusters the display.

```{r, width=6, height=1.6}
bind_rows(ud_combined, .id = "bootstrap") %>% 
  mutate(bootstrap = factor(bootstrap, levels = c("parametric", "nonparametric", "compromise"))) %>%
  plot_overlay_truth(100)
```
