---
title: "Generate Simulation Images"
output: html_document
params:
  output_dir: "../../data/raw_data/"
---

This notebook simulates images according to the log Cox Matern Process discussed
in Section 3.2 and plotted in Figure 3. Note that these data are already
available in the `stability_data_sim.tar.gz` archive within `raw_data.tar.gz`
linked in Supplementary Section A. However, if the goal was to regenerate those
input data archives, this script could be used.

Generating each image takes a few seconds, and we need many training images, so
we've parallized generation across cores. This is why we're loaded the
`doParallel`, `foreach` and `parallel` packages.

```{r}
library(MSLF)
library(doParallel)
library(foreach)
library(parallel)
library(reticulate)
library(tidyverse)
np <- reticulate::import("numpy")
```

This is a specification of all the parameters given in Table 1.

```{r}
x <- expand.grid(seq(0.1, 1, 0.05), seq(0.1, 1, 0.05))
n_original <- 100 # number of points
beta_r <- c(-0.5, 0, 0.5)
nu_r <- 1 # smoothness of the relative intensity functions (higher is smoother)
alpha_r <- 0.5 # bandwidth of the relative processes (higher is smoother)
nu <- 2 # smoothness of the overall process
alpha <- 2 # bandwidth of the overall process
tau <- 1.5 # temperature for the marking (0 -> uniform, inf -> hard assignment)
lambdas <- c(20, 50, 100) # rate parameters in gamma controlling size
```

The block below shows an example image.

```{r, fig.width = 7, fig.height = 3}
set.seed(10)
result <- sim_wrapper(x, n_original, nu, alpha, beta_r, nu_r, alpha_r, tau, lambdas)
plot_matern(result)
```

# Simulating Images

Next we relate the parameters underlyign each image to the outcome $y$
associated with each image. We made the following assumptions,

* More mixing between cell types is associated with better outcomes
* Less uniform distribution in cell densities is associated with worse outcomes.
* More of the first cell type is better for outcomes, more of the other two is bad.
* larger of the first cell type is better for outcomes

so that the signs are in the intuitive direction and so that no one group of
terms dominates.

We generate the parameters according to a few different distributions, depending
on whether they are allowed to be negative.

```{r}
n_images <- 1e4
n_original <- runif(n_images, 50, 1000)
beta_r <- 0.3 * runif(n_images * 3, -.5, .5) %>%
  matrix(ncol = 3)
nu_r <- runif(n_images, 0, 3)
alpha_r <- runif(n_images, 0, 3)
nu <- runif(n_images, 0, 8) # overall process is (usually) smoother
alpha <- runif(n_images, 0, 8)
tau <- runif(n_images, 0, 3)
lambdas <- runif(n_images * 3, 100, 500) %>%
  matrix(ncol = 3)

# simulate the response
X <- scale(cbind(n_original, beta_r, nu_r, alpha_r, nu, alpha, tau, lambdas[, 1]))
colnames(X) <- c("N", paste0("beta_", 1:ncol(beta_r)), "nu_r", "alpha_r", "nu", "alpha", "tau", "lambda_1")
y <- X %*% c(.5, 1, -1, -1, -.5, -.5, -.5, -.5, .5, 1) / sqrt(ncol(X))
y <- y - mean(y)
```

This block plots two examples each for three very different $y$ ranges. Each
component in `plot_list` is one of the panels appearing in Figure 3.

```{r, eval = FALSE}
plot_ix <- order(y)[c(1:2, c(4998, 5004), c(9998, 10000))]
plot_list <- list()
raster_list <- list()

for (i in seq_along(plot_ix)) {
  j <- plot_ix[i]
  res <- sim_wrapper(x, n_original[j], nu[j], alpha[j], beta_r[j, ], nu_r[j], alpha_r[j], tau[j], lambdas[j, ])
  raster_list[[i]] <- plot_raster(make_raster(res$marks, 3))
  plot_list[[i]] <- plot_matern(res)
}
```

This combines all the figures in `plot_list` and `raster_list` above to create
Figure 3.

```{r, eval = FALSE, fig.height = 5, fig.width = 10}
library(gridExtra)
grid.arrange(
  plot_list[[1]], raster_list[[1]],
  plot_list[[2]], raster_list[[2]],
  plot_list[[3]], raster_list[[3]],
  plot_list[[4]], raster_list[[4]],
  plot_list[[5]], raster_list[[5]],
  plot_list[[6]], raster_list[[6]],
  ncol = 4,
  widths = c(4, 1, 4, 1)
)
```

We are almost ready to simulate all the images. The block below saves the
directory to which all the simulated outputs will be saved.

```{r}
tiles_dir <- file.path(params$output_dir, "tiles")
dir.create(tiles_dir)
```

Finally, we generate the images. The `Xy.csv` file saved at the very end saves
all the raw parameter and response values associated with each generated image.

```{r}
# can convert foreach into a standard for loop if no parallel backend
cl <- makeCluster(12)
registerDoParallel(cl)

pkgs <- c("inference", "dplyr", "MASS", "stringr", "sf", "raster", "reticulate")
paths <- foreach(j = seq_len(n_images), .combine = c, .packages = pkgs) %dopar% {
  # generate the shapefile representing cells
  result <- sim_wrapper(x, n_original[j], nu[j], alpha[j], beta_r[j, ], nu_r[j],
                        alpha_r[j], tau[j], lambdas[j, ])
  marks <- result$marks %>%
    mutate(variable = paste0("X", mark))
  
  # convert to raster and save as numpy
  path <- sprintf(file.path(tiles_dir, "image-%s.npy"), str_pad(j, 4, "left", 0))
  r <- make_raster(marks, 3)
  np$save(path, as.array(r))
  return(path)
}
stopCluster(cl)

write_csv(data.frame(path = paths, X = X, y = y), file.path(params$output_dir, "Xy.csv"))
```
